---
title: "NPBin"
author: "Anthony Aylward"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We will analyze a subset of one of the sample dataset for illustration
purposes.
```{r}
library(npbin)

minimum_coverage <- 5 # minimum total coverage allowed
n_cores <- 47 # the number of cores to be used, can ONLY be 1 if run on Windows.

dt.ct <- atac
colnames(dt.ct)
```

for illustration purpose, keep only the 2000 points, remove this line will
end up with analyzing the whole data set. It could be slow if only one core
is used.
```{r}
dt.ct <- dt.ct[1:2000, ]
```

NPBin
```{r}
n_breaks <- 11 # number of breaks
spline_order <- 4 # order of splines
breaks <- seq(0, 1, length.out = n_breaks)
pi_init <- hist(
  dt.ct[, p_hat],
  breaks = seq(0, 1, length.out = n_breaks + spline_order - 3),
  plot = FALSE
)[["density"]] # initialized the weights using the histogram of p_hat
```

estimate the overall model
```{r}
overall_model_estimate <- emBinBspl(
  dt.ct[, xm],
  dt.ct[, m],
  breaks = breaks,
  k = spline_order,
  pi.init = pi_init,
  ncores = n_cores,
  err.max = 1e-3,
  iter.max = 200
)  
```

estimate the null model
```{r}
null_model_estimate <- estNull(
  dt.ct[, xm],
  dt.ct[, m],
  overall_model_estimate,
  init = NULL,
  iter.max = 200,
  ncores = n_cores,
  ub = rep(log(1e4), 2),
  err.max = 1e-4
)
dt.ct[,
  fnp := null_model_estimate[["f"]]
][,
  f0np := null_model_estimate[["f0"]]
][,
  locfdrnp := null_model_estimate[["locfdr"]]
][,
  fdrnp := locfdr2FDR(locfdrnp)
][,
  ranknp := rank(locfdrnp, ties.method = "max")
]
names(null_model_estimate)
```

```{r}
null_model_estimate$coef.null
```

Empirical Bayes test using p_hat
```{r}
pct0 <- 0.45         
empirical_bayes_beta_hat <- ebBeta(
  dt.ct[, xm],
  dt.ct[, m],
  dt.ct[, p_hat],
  breaks = breaks,
  k = spline_order,
  pi.init = pi_init,
  pct0 = pct0,
  init = NULL,
  iter.max = 200,
  err.max = 1e-4,
  ncores = n_cores
)
dt.ct[,
  fhat := empirical_bayes_beta_hat[["f"]]
][,
  f0hat := empirical_bayes_beta_hat[["f0"]]
][,
  locfdrhat := empirical_bayes_beta_hat[["locfdr"]]
][,
  fdrhat := locfdr2FDR(locfdrhat)
][,
  rankhat := rank(locfdrhat, ties.method = "max")
]
names(empirical_bayes_beta_hat)
```

null parameters of EBE
```{r}
empirical_bayes_beta_hat[["coef.null"]]
```

Binomial test
```{r}
p_binomial <- sapply(
  1:n,
  function(y) binom.test(dt.ct[y, xm], dt.ct[y, m])[["p.value"]]
)
dt.ct[,
  pvbin := p_binomial
][,
  fdrbin := p.adjust(pvbin, method = "BH")
][,
  rankbin := rank(pvbin, ties.method = "max")
]
```

Evaluate the results using motifs

number of potential TP defined based on motif
```{r}
dt.ct[, sum(potential_TP)]
```

number of potential FP defined based on motif
```{r}
dt.ct[, sum(potential_FP)]
```

find the number of TP and FP in top ranked SNPs
```{r}
dt.ct[,
  tpnp := rank2nhit(ranknp ,potential_TP)
][,
  fpnp := rank2nhit(ranknp, potential_FP)
]
dt.ct[,
  tp_hat := rank2nhit(rankhat, potential_TP)
][,
  fp_hat := rank2nhit(rankhat, potential_FP)
]
dt.ct[,
  tpbin := rank2nhit(rankbin, potential_TP)
][,
  fpbin := rank2nhit(rankbin, potential_FP)
]
```

plot the accuracy measure as in the main paper. 
We presented a zoom-in version in the main paper to the top 20%, 
because usually there are not many ALI SNPs.
Note that the default of the demo only select a subset of the data for
illustration purposes.
Thus the figure may not an exact replica of the one in the paper.
To reproduce the results in the paper, please use the whole dataset
```{r, fig.show='hold'}
cbfpalette <- c(
  "#D55E00",
  "#0072B2",
  "#CC79A7",
  "#009E73",
  "#E69F00",
  "#56B4E9",
  "#F0E442"
)
plotidac <- c(' NPB','EBE','Binom')
```