---
title: "NPBin"
author: "Anthony Aylward"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We will analyze a subset of one of the sample dataset for illustration
purposes.
```{r shape_dnase_import_packages}
library(parallel)
library(npbin)
library(data.table)

minimum_coverage <- 5 # minimum total coverage allowed
n_cores <- detectCores() # the number of cores to be used, can ONLY be 1 if run on Windows.

dt <- dnase
colnames(dt)
```

```{r shape_dnase_initialize_data}
dt.ct <- data.table(dt)[m >= minimum_coverage, ]
```

for illustration purpose, keep only the 2000 points, remove this line will
end up with analyzing the whole data set. It could be slow if only one core
is used.
```{r shape_dnase_subset_data}
dt.ct <- dt.ct[1:2000, ]
dt.ct[, p_hat:=xm / m]
n <- nrow(dt.ct)
```

NPBin
```{r shape_dnase_initialize_weights}
n_breaks <- 11 # number of breaks
spline_order <- 4 # order of splines
breaks <- seq(0, 1, length.out = n_breaks)
pi_init <- initialize_weights(
  dt.ct,
  n_breaks = n_breaks,
  spline_order = spline_order,
  plot = TRUE
) # initialized the weights using the histogram of p_hat
```

estimate the overall model
```{r shape_dnase_estimate_overall}
overall_model_estimate <- emBinBspl(
  dt.ct[, xm],
  dt.ct[, m],
  breaks = breaks,
  k = spline_order,
  pi.init = pi_init,
  ncores = n_cores,
  err.max = 1e-3,
  iter.max = 200
)
names(overall_model_estimate)
```

estimate the null model
```{r shape_dnase_estimate_null}
null_model_estimate <- estNull(
  dt.ct[, xm],
  dt.ct[, m],
  overall_model_estimate,
  init = NULL,
  iter.max = 200,
  ncores = n_cores,
  ub = rep(log(1e4), 2),
  err.max = 1e-4
)
names(null_model_estimate)
```

```{r shape_dnase_null_parameters}
null_model_estimate[["coef.null"]]
```

Calculate the [overdispersion/correlation parameter](https://en.wikipedia.org/wiki/Beta-binomial_distribution#Moments_and_properties)

```{r shape_dnase_overdispersion}
1 / sum(
  null_model_estimate[["coef.null"]][["shape1"]],
  null_model_estimate[["coef.null"]][["shape2"]],
  1
)
```

Plot the estimated null distribution

```{r shape_ctcf_plot_null}
plot_estimated_null(
  dt.ct[, p_hat],
  null_model_estimate[["coef.null"]][["shape1"]],
  null_model_estimate[["coef.null"]][["shape2"]]
)
```
